Kronos \& Krios: Evidence for accretion of a massive, rocky planetary system in a comoving pair of solar-type stars

### 2017-09-10 paper/ms.tex
discussion and summary

We discuss the possible origins of the peculiar abundance differences of
\bizarreone-\sunanalog.
We first discuss the ages and coevality of the stars in this pair, and consider
both possibilities in which the two stars are or are not coeval.
Our favored scenario is discussed in the last subsection,
\sectionname~\ref{sub:accretion}.

@@

{\bf
The surface lithium abundance in a Sun-like star decreases with its age due to
mixing induced by convection or rotation, which brings the lithium into the
interior ($T>2.5 \times 10^{6}$~K) where it will be destroyed by proton capture
burning.
Thus, surface lithium abundance can be an indicator of stellar ages,
especially whether the star is very young ($\lesssim 1$~Gyr).
The absolute $\elem{Li}$ abundance of solar-type stars also correlates steeply
with the effective temperature (e.g., \citealt{Chen2001,2012ApJ...756...46R}).
Generally, cooler stars with larger convective envelope have lower \elem{Li}
abundance.
The absolute $\elem{Li}$ abundance of $2.25$~dex for \sunanalog\
is typical for its \teff.
On the other hand, the lithium abundance ($A(\elem{Li}) = 2.75$) of \bizarreone,
which has lower \teff\ than \sunanalog, is not only higher than that of
\sunanalog\ but also much higher compared to other field stars of similar \teff.
Given the overall higher metal abundances and the peculiar abundance patterns
in \bizarreone, it is unclear, however, whether this higher $\elem{Li}$
abundance means a younger age or something else.
For example, \citealt{Casey:2016aa} attributes the presence of $\elem{Li}$-rich
red giant stars to the engulfment of substellar companions such as gas giant
planets or brown dwarfs which may replenish $\elem{Li}$.
}

@@

Another possibility that two coeval stars may end up with different surface
abundances is accretion of planetary material after birth.
In a multi-planet system, dynamical instabilities triggered by planet-planet
scattering\cite{1996Sci...274..954R,1996Natur.384..619W} or encounters with a
field star\cite{Malmberg:2011aa} can lead to planet ejection or accretion.
Indeed, it is an important goal of many exoplanet studies
to detect chemical signatures of planet formation or accretion,
distinguish them from Galactic chemical evolution, and
connect them to theories of evolution of planetary systems.
One approach that is free from confusion with Galactic chemical evolution
is to compare two almost identical stars in a wide binary system.
Assuming that the component stars were born together with identical
initial composition, we may see a difference in their surface abundances
if the two stars then accreted different amounts of planetary material.
The resulting abundance difference may depend on the condensation
temperatures of elements in the protoplanetary disks from which the accreted
planets formed, as their compositions depend on the radial temperature gradient
in the disk.
%TODO: cautions on condensation temperature
% 1. it is just a proxy
% 2. it is **equlibrium** condensation temperature

@@

How much mass of rocky material is needed to explain an increment of
$\approx 0.2$~dex?
We carry out simple toy calculations of expected $\Delta\elemH{X}$
in a Sun-like star's atmosphere by adding certain mass of bulk Earth composition
under these simplifying assumptions:
\begin{itemize}
  \item The material added is instantly and completely mixed.
  \item The atmospheric composition that we measure is identical throughout
    the star's radiative and convective zone.
  \item The surface abundance of the star has been altered only by the
    accretion event(s).
\end{itemize}

We take the solar abundances, $\elemH{X}$, of element \elem{X}
(\citealt{Asplund:2009aa}) which can be converted to mass fraction as
\begin{equation}
  f_{X,\mathrm{photo}} = \frac{10^{\elemH{X}} m_X}{\Sigma_X 10^{\elemH{X}} m_X}
\end{equation}
where $m_X$ is the mass of each element in e.g., atomic mass unit.
Assuming that the accreted material has a total mass $M_\mathrm{acc}$, and the
mass fraction in each element $f_{X,\mathrm{acc}}$,
the abundance difference is
\begin{equation}
  \Delta\elemH{X} = \log_{10} \frac{f_{X,\mathrm{photo}}\,f_\mathrm{CZ}\,M_\mathrm{star}
    + f_{X,\mathrm{acc}}\,M_\mathrm{acc}}
    {f_{X,\mathrm{photo}}\,f_\mathrm{CZ}\,M_\mathrm{star}}
\end{equation}
where $f_\mathrm{CZ}$ is the fraction of the star's mass in the convective envelope.
We assume $f_\mathrm{CZ} = 0.02$ (\citealt{2013ApJ...776...87S}),
and take the composition of bulk Earth from a chondritic model of the Earth
(\citealt{2003TrGeo...2..547M}).
Similar calculations have been performed by e.g., \citet{Chambers:2010aa},
\citet{Mack:2014aa,Mack:2016aa}.

Figure~\ref{fig:toycalc} shows the expected change of surface abundances of
metals in a Sun-like star after \maccreted\ of material with composition of
bulk Earth is added.
A volatility trend that more volatile (low \Tcondens) elements are more
depleted in the Earth relative to CI or other carbonaceous chondrites
has long been known (\citealt{mcdonough2001composition}).
This trend is presumed to be closely related to the formation of terrestrial
planets and, in particular to the radial temperature gradient in a
protoplanetary disk.
The trend resulting from adding \maccreted\ of bulk Earth
provides an overall good match to the observed $\Delta\elemH{X}$,
suggesting that the refractory-enhanced star, \bizarreone\,
accreted \maccreted\ more of rocky planetary material than \sunanalog.

What about \elem{Li}?
The element \elem{Li} is worth special attention in the context of the
accretion scenario.
Because Li is present in either carbonaceous chondrites or bulk Earth with a
concentration of $1-1.5$~ppm in mass (\citealt{2003TrGeo...2..547M}), but
is depleted quickly within the first couple of Gyrs on the surface of a
Sun-like star (\askjmb{CITE?}), accretion of either material at later times
will significantly replenish the lithium on the star's surface.
For the present-day Sun ($A(\elem{Li}) = 1.05$), the accretion of \maccreted\ of
bulk Earth-like material would result in $\Delta\elemH{Li} \approx 1.65$~dex
(see the inset of \figname~\ref{fig:toycalc}).
This closely matches what we find: the \elem{Li} abundance of \bizarreone\ is
$A(\elem{Li}) = 2.75$ (Table~\ref{tab:kk}, \citealt{jmlithium})
approximately $1.7$~dex higher than the solar value.

@@

We report and discuss the discovery of a comoving pair of bright
solar-type stars HD~240430 and HD~240429 (G0 and G2) with very different
metallicities ($\Delta\feh \approx 0.2$~dex), and condensation temperature
(\Tcondens)-dependent abundance differences.
The more metal-rich of the two stars, HD~240430 (\bizarreone), shows enhancement in
all ten elements with $\Tcondens > 1200$~K including \elem{Fe}, while under
enhanced in the five elements, \elem{C}, \elem{N}, \elem{O}, \elem{Na}, and
\elem{Mn} with $\Tcondens < 1200$~K relative to HD~240429 (\sunanalog).
It also has an anomalously high surface \elem{Li} abundance for its age of
$\sim 4$~Gyr, and its effective temperature very close to that of the Sun.
We consider that the comoving pair may have formed from two stars of different
birth origins in a exchange scattering event
(\sectionname~\ref{sub:exchange_scattering}), or that there may be chemical
inhomogeneity in the birth cloud
(\sectionname~\ref{sub:chemical_inhomogeneity_in_star_formation}) to find
both unlikely.

In order to explain the $\Tcondens$-dependent enhancement and high \elem{Li}
abundance of \bizarreone, we consider accretion of planetary material as the
cause (\sectionname~\ref{sub:accretion}).
We argue that an accretion of \maccreted\ of bulk Earth composition to
\bizarreone\ after its thin convective zone is in place can explain the
enhancement in both refractory elements and lithium.
What triggered the planet engulfment in the two comoving stars remains unclear.
One possibility is that a fly-by interaction with a field star could have
triggered eccentricity excitation of outer planets around each star, which may
have propagated inward through planet-planet scattering, leading to the
accretion of inner rocky planets (\citealt{Zakamska:2004aa,Malmberg:2011aa}).
If this is the case, there may be surviving, highly-eccentric giant planets.
The two stars have not been included in any publicly released data from planet
search programs, however, outer eccentric planets are potentially detectable
with future data releases of the \gaia\ mission.

>>>>

We discuss the possible origins of the peculiar abundance differences of
\bizarreone\ \& \sunanalog.
We first discuss the ages and coevality of the stars in this pair, and consider
both possibilities in which the two stars are or are not coeval.
Our favored scenario is discussed in the last subsection,
\sectionname~\ref{sub:accretion}.

@@

{\bf
The surface lithium abundance in a Sun-like star decreases with its age due to
mixing induced by convection or rotation, which brings the lithium into the
interior ($T>2.5 \times 10^{6}$~K) where it will be destroyed by proton capture
burning.
Thus, surface lithium abundance can be an indicator of stellar ages,
especially whether the star is very young ($\lesssim 1$~Gyr).
The absolute $\elem{Li}$ abundance of solar-type stars also correlates steeply
with the effective temperature (e.g., \citealt{Chen2001,2012ApJ...756...46R}).
Generally, cooler stars with larger convective envelope have lower \elem{Li}
abundances.
The absolute $\elem{Li}$ abundance of $2.25$~dex for \sunanalog\
is typical for its \teff.
On the other hand, the lithium abundance ($A(\elem{Li}) = 2.75$) of \bizarreone,
which has lower \teff\ than \sunanalog, is not only higher than that of
\sunanalog\ but also much higher compared to other field stars of similar \teff.
Given the overall higher metal abundances and the peculiar abundance patterns
in \bizarreone, it is unclear, however, whether this higher $\elem{Li}$
abundance means a younger age or something else.
For example, \citealt{Casey:2016aa} attributes the presence of $\elem{Li}$-rich
red giant stars to the engulfment of substellar companions such as gas giant
planets or brown dwarfs which may replenish $\elem{Li}$.
}

@@

Another possibility that two coeval stars may end up with different surface
abundances is accretion of planetary material after birth.
In a multi-planet system, dynamical instabilities triggered by planet-planet
scattering (\citealt{1996Sci...274..954R,1996Natur.384..619W}) or encounters with a
field star (\citealt{Malmberg:2011aa}) can lead to planet ejection or accretion.
Indeed, it is an important goal of many exoplanet studies
to detect chemical signatures of planet formation or accretion,
distinguish them from Galactic chemical evolution, and
connect them to theories of evolution of planetary systems.
One approach that is free from confusion with Galactic chemical evolution
is to compare two almost identical stars in a wide binary system.
Assuming that the component stars were born together with identical
initial composition, we may see a difference in their surface abundances
if the two stars then accreted different amounts of planetary material.
The resulting abundance difference may depend on the condensation
temperatures of elements in the protoplanetary disks from which the accreted
planets formed, as their compositions depend on the radial temperature gradient
in the disk.
%TODO: cautions on condensation temperature
% 1. it is just a proxy
% 2. it is **equlibrium** condensation temperature

@@

How much mass of rocky material is needed to explain an increment of
$\approx 0.2$~dex?
We carry out simple toy calculations of the expected $\Delta\elemH{X}$
in a Sun-like star's atmosphere by adding a certain mass of bulk Earth composition
under these simplifying assumptions:
\begin{itemize}
  \item The material added is instantly and completely mixed through the star's convective zone.
  \item The atmospheric composition that we measure is identical throughout
    the star's radiative and convective zone.
  \item The surface abundance of the star has been altered only by the
    accretion event(s).
\end{itemize}

We take the solar abundances, $\elemH{X}$, of element \elem{X}
(\citealt{Asplund:2009aa}) which can be converted to mass fraction as
\begin{equation}
  f_{X,\mathrm{photo}} = \frac{10^{\elemH{X}} \, m_X}{\Sigma_X 10^{\elemH{X}} \, m_X}
\end{equation}
where $m_X$ is the mass of each element in, e.g., atomic mass unit.
Assuming that the accreted material has a total mass $M_\mathrm{acc}$, and the
mass fraction in each element $f_{X,\mathrm{acc}}$,
the abundance difference is
\begin{equation}
  \Delta\elemH{X} = \log_{10} \frac{f_{X,\mathrm{photo}}\,f_\mathrm{CZ}\,M_\mathrm{star}
    + f_{X,\mathrm{acc}}\,M_\mathrm{acc}}
    {f_{X,\mathrm{photo}}\,f_\mathrm{CZ}\,M_\mathrm{star}}
\end{equation}
where $f_\mathrm{CZ}$ is the fraction of the star's mass in the convective envelope.
We assume $f_\mathrm{CZ} = 0.02$ (\citealt{2013ApJ...776...87S}),
and take the composition of bulk Earth from a chondritic model of the Earth
(\citealt{2003TrGeo...2..547M}).
Similar calculations have been performed by, e.g., \citet{Chambers:2010aa},
\citet{Mack:2014aa,Mack:2016aa}.

Figure~\ref{fig:toycalc} shows the expected change of surface abundances of
metals in a Sun-like star after \maccreted\ of material with composition of
bulk Earth is added.
A volatility trend such that more volatile (low \Tcondens) elements are more
depleted in the Earth relative to CI or other carbonaceous chondrites
has long been known (\citealt{mcdonough2001composition}).
This trend is presumed to be closely related to the formation of terrestrial
planets and, in particular, to the radial temperature gradient in a
protoplanetary disk.
The trend resulting from adding \maccreted\ of bulk Earth
provides an overall good match to the observed $\Delta\elemH{X}$,
suggesting that the refractory-enhanced star, \bizarreone\,
accreted \maccreted\ more of rocky planetary material than \sunanalog.

What about \elem{Li}?
The element \elem{Li} is worth special attention in the context of the
accretion scenario.
Because Li is present in either carbonaceous chondrites or bulk Earth with a
concentration of $1-1.5$~ppm in mass (\citealt{2003TrGeo...2..547M}), but
is depleted quickly within a few gigayears on the surface of a
Sun-like star (\askjmb{CITE?}), accretion of either material at later times
will significantly replenish the lithium on the star's surface.
For the present-day Sun ($A(\elem{Li}) = 1.05$), the accretion of \maccreted\ of
bulk Earth-like material would result in $\Delta\elemH{Li} \approx 1.65$~dex
(see the inset of \figname~\ref{fig:toycalc}).
This closely matches what we find: the \elem{Li} abundance of \bizarreone\ is
$A(\elem{Li}) = 2.75$ (Table~\ref{tab:kk}, \citealt{jmlithium})
approximately $1.7$~dex higher than the solar value.

@@

We report and discuss the discovery of a comoving pair of bright
solar-type stars HD~240430 and HD~240429 (G0 and G2) with very different
metallicities ($\Delta\feh \approx 0.2$~dex), and condensation temperature
(\Tcondens)-dependent abundance differences.
The more metal-rich of the two stars, HD~240430 (\bizarreone), shows enhancement in
all ten elements with $\Tcondens > 1200$~K including \elem{Fe}, while
under-enhanced in the five elements, \elem{C}, \elem{N}, \elem{O}, \elem{Na}, and
\elem{Mn} with $\Tcondens < 1200$~K relative to HD~240429 (\sunanalog).
It also has an anomalously high surface \elem{Li} abundance for its age of
$\sim 4$~Gyr, and its effective temperature very close to that of the Sun.
We consider that the comoving pair may have formed from two stars of different
birth origins in a exchange scattering event
(\sectionname~\ref{sub:exchange_scattering}), or that there may be chemical
inhomogeneity in the birth cloud
(\sectionname~\ref{sub:chemical_inhomogeneity_in_star_formation}) to find
both unlikely.

In order to explain the $\Tcondens$-dependent enhancement and high \elem{Li}
abundance of \bizarreone, we consider the accretion of planetary material as the
most plausible cause (\sectionname~\ref{sub:accretion}).
We argue that an accretion of \maccreted\ of bulk Earth composition to
\bizarreone\ after its thin convective zone is in place can explain the
enhancement in both refractory elements and lithium.
What triggered the planet engulfment in the two comoving stars remains unclear.
One possibility is that a fly-by interaction with a field star could have
triggered eccentricity excitation of outer planets around each star, which may
have propagated inward through planet-planet scattering, leading to the
accretion of inner rocky planets (\citealt{Zakamska:2004aa,Malmberg:2011aa}).
If this is the case, there may be surviving, highly-eccentric giant planets.
The two stars have not been included in any publicly released data from planet
search programs, however, outer eccentric planets are potentially detectable
with future data releases of the \gaia\ mission.
